---
title: "Grouping Games Through Clustering"
author: "John Lim"
---

```{r}
#| label: setup
#| include: false

# improve digit and NA display 
options(scipen = 1, knitr.kable.NA = '')

# load packages
library(tidyverse)
library(readr)
library(purrr)
library(broom)
library(ggplot2)
library(ggrepel)
library(GGally)
library(kableExtra)
```

```{r}
#| label: read-files

# read data for clustering
game_data <- readRDS("data/processed/clean_games.rds")
```

To address the question about which video games are similar (and how), I wanted to use unsupervised learning techniques. I will use *k*-means clustering as a way to assign video games to distinct groups, which may uncover similarities between the games. 

For this analysis, I chose 9 quantitative variables to cluster the games by: the age required to play, the price, the number of DLCs (Downloadable Content), the metacritic score, the number of game achievements, the number of recommendations, the median playtime since the game's release, the highest CCU (Concurrent User) count, and the net review ratio statistic. Since the variables on different scales (e.g. 0 to 100, -1 to 1, and 0 to 1,000,000+), it is important to standardize the variables (i.e. level the playing field). This way, the locations of the centroids will not be dominated by its relation to one variable over another when using *k*-means clustering.

```{r}
#| label: cluster-prep

game_standardized <- game_data |>
  # get quantitative variables that I want to use for clustering
  select(required_age, price, dlc_count, metacritic_score,
         achievements, recommendations,
         median_playtime_forever, peak_ccu, net_review_ratio) |>
  # make sure observations are complete
  drop_na() |>
  # standardize variables for clustering
  mutate(across(everything(),
                ~ (.x - mean(.x)) / sd(.x),
                .names = "{.col}_z")) |>
  select(ends_with("_z"))
```

To determine the optimal number of clusters, I first consulted an elbow plot. We want to look at when the within-cluster variance is relatively small, without creating too many clusters. However, with too many clusters, it may become too difficult to interpret the results or for the results to hold any significant meaning. Therefore, based on the elbow plot, 5 clusters appears to be reasonably optimal as it is where the graph "bends" without containing too many clusters.

```{r}
#| label: elbow-plot
#| fig-cap: "Elbow Plot Indicates A Bend at 5 Clusters"

# set seed for reproducibility
set.seed(42069777)

# iterate through clustering algorithm for 10 different values of k
elbow_plot <- tibble(k = 1:10) |>
  mutate(
    kmeans_results = purrr::map(k, ~ kmeans(game_standardized, .x, nstart=25)),
    glanced = purrr::map(kmeans_results, glance)) |>
  unnest(cols = c(glanced))

# construct elbow plot
ggplot(elbow_plot, aes(x = k, y = tot.withinss)) +
  geom_point() + 
  geom_line() +
  geom_point(data = (elbow_plot |> filter(k == 5)), color = "red", size = 3) +
  scale_x_continuous(breaks = 1:10) +
  labs(x = "Number of Clusters (k)", 
       y = "Total Within-Cluster Sum of Squares",
       title = "Elbow Plot with Standardized Variables") +
  theme_classic()
```

```{r}
#| label: fit-cluster
#| tbl-cap: "Locations of Centroids based on Standardized Variables"
#| tbl-cap-location: top

# for reproducibility
set.seed(42069777)

# save 5 cluster object
games_kmeans5 <- game_standardized |>
  kmeans(centers = 5, nstart = 25)

# augment original dataset to include which cluster each game belongs to 
games_c5 <- augment(games_kmeans5, game_data) |>
  rename(cluster_standardized = .cluster) |>
  select(name, release_date, required_age, price,
         dlc_count, metacritic_score, achievements, recommendations,
         median_playtime_forever, peak_ccu, net_review_ratio,
         cluster_standardized)

# present table that shows where these centroids are located
games_kmeans5$centers |>
  as_tibble() |>
  mutate(Cluster = row_number(), .before = 1) |>
 kable(booktabs = TRUE,
       col.names = c("Cluster","Required Age", "Price", "DLC Count",
                      "Metacritic Score", "Achievements", "Recommendations",
                      "Median Playtime Forever", "Peak CCU",
                      "Net Review Ratio"),
       caption = "Note: Each Quantitative Variable is Standardized",
       digits = 3) |>
  kable_styling()

```

It appears that our clusters are grouped by the following characteristics:

1. Higher DLC counts, extremely high number of achievements, and lower ratings.

2. Generally average values across the variables except for lower metacritic scores and high ratings.

3. Highest age requirements, highest prices, higher metacritic scores, high number of recommendations, and the highest CCU counts.

4. Lowest median playtimes and lowest ratings.

5. Higher prices, highest DLC counts, highest metacritic scores, and highest median playtimes.

Therefore, Cluster 1 contains games that may cater towards completionist gamers, although lacking in ratings. Cluster 2 contains games that are not approved by metacritics but supported and favored by players. Cluster 3 contains games that are catered towards older audiences and also widely shared. Cluster 4 contains games that plain out suck; players do not play these games for long and give bad ratings. Finally, Cluster 5 contains expensive games with a lot of content for committed players.

```{r}
#| label: game-example
#| tbl-cap: "Table Featuring Games From Cluster 5"

# # showcase games in cluster 1
# games_c5 |>
#   filter(cluster_standardized == 1) |>
#   arrange(desc(achievements), desc(dlc_count), net_review_ratio) |>
#   head() |>
#   kable(booktabs = TRUE,
#        col.names = c("Name","Release Date","Required Age", "Price",
#                      "DLC Count", "Metacritic Score", "Achievements",
#                      "Recommendations", "Median Playtime Forever", "Peak CCU",
#                      "Net Review Ratio", "Cluster"),
#        digits = 3) |>
#   kable_styling()
# 
# # showcase games in cluster 2
# games_c5 |>
#   filter(cluster_standardized == 2) |>
#   arrange(desc(net_review_ratio), metacritic_score) |>
#   head() |>
#   kable(booktabs = TRUE,
#        col.names = c("Name","Release Date","Required Age", "Price",
#                      "DLC Count", "Metacritic Score", "Achievements",
#                      "Recommendations", "Median Playtime Forever", "Peak CCU",
#                      "Net Review Ratio", "Cluster"),
#        digits = 3) |>
#   kable_styling()
# 
# # showcase games in cluster 3
# games_c5 |>
#   filter(cluster_standardized == 3) |>
#   arrange(desc(required_age), desc(price), desc(recommendations),
#           desc(peak_ccu), desc(metacritic_score)) |>
#   head() |>
#   kable(booktabs = TRUE,
#        col.names = c("Name","Release Date","Required Age", "Price",
#                      "DLC Count", "Metacritic Score", "Achievements",
#                      "Recommendations", "Median Playtime Forever", "Peak CCU",
#                      "Net Review Ratio", "Cluster"),
#        digits = 3) |>
#   kable_styling()
# 
# # showcase games in cluster 4
# games_c5 |>
#   filter(cluster_standardized == 4) |>
#   arrange(desc(net_review_ratio), desc(median_playtime_forever)) |>
#   head() |>
#   kable(booktabs = TRUE,
#        col.names = c("Name","Release Date","Required Age", "Price",
#                      "DLC Count", "Metacritic Score", "Achievements",
#                      "Recommendations", "Median Playtime Forever", "Peak CCU",
#                      "Net Review Ratio", "Cluster"),
#        digits = 3) |>
#   kable_styling()

# showcase games in cluster 5
games_c5 |>
  filter(cluster_standardized == 5) |>
  arrange(desc(metacritic_score), desc(median_playtime_forever), desc(dlc_count), desc(price)) |>
  head() |>
  kable(booktabs = TRUE,
       col.names = c("Name","Release Date","Required Age", "Price",
                     "DLC Count", "Metacritic Score", "Achievements",
                     "Recommendations", "Median Playtime Forever", "Peak CCU",
                     "Net Review Ratio", "Cluster"),
       digits = 3) |>
  kable_styling()

```

From the table, we can see that *Portal 2*, *Sid Meier's Civilization® IV*, and *Tetris Effect: Connected* are similar as they have high Metacritic Scores and high median playtimes.